# Reproducible Research: Peer Assessment 1


================================================


## Loading and preprocessing the data


We assume that the activity.csv file has already been unzipped in the workdirectory.Lets check what is the number of rows with completed obseravtions.

```{r, results="asis"}
setwd("C:/Users/x/RepData_PeerAssessment1")
act<-read.csv("activity.csv")
head(act)
dim (act)
cc<-complete.cases(act)
sum(cc) #This is the number of complete observations
act1<-act[cc,] #This is a subset with completed observations only
dim (act1)
```


## What is mean total number of steps taken per day?

```{r, results="asis"}
days<-as.Date(act1$date)
total_steps<-tapply(act1$steps,days,sum, na.rm=TRUE)
length(total_steps)
hist(total_steps, main="Histogram total number of steps per day")
mean_steps<-tapply(act1$steps, days, mean, na.rm=TRUE) #This is the mean number of steps per day
mean_steps
plot(mean_steps)
median_steps<-tapply(act1$steps, days, median, na.rm=TRUE) #This is the median number of steps per day
median_steps
plot(median_steps)
```

## What is the average daily activity pattern?

```{r, results="asis"}
mean_steps_int<-tapply(act1$steps,act1$interval,mean,na.rm=TRUE)
length(mean_steps_int)
plot(mean_steps_int~levels(factor(act1$interval)), type="l")
grep(max(mean_steps_int), mean_steps_int)
levels(factor(act1$interval))[[104]]#This is the time interval with maximal activity
```

## Imputing missing values

```{r, results="asis"}
means_int<-tapply(act$steps, act$interval, mean, na.rm=TRUE)
length(means_int)
ndays<-nrow(act)/length(means_int)
seq<-rep(means_int, times=ndays)
act_imput<-cbind(act,seq)
fac<-is.na(act$steps)
sub1<-act_imput[fac,]
sub2<-act_imput[!fac,]
sub1[,1]<-sub1[,4]
act_final<-rbind(sub1,sub2)
table(is.na(act_final))
act_final<-act_final[,1:3]
days1<-as.Date(act_final$date)
head(act_final)
length(table(days1))## Obviously we have 8 more days with complete data.
mean_steps_final<-tapply(act_final$steps, days1, mean)
mean_steps_final #Means did not change because measurements for entire day were excluded
plot(mean_steps_final)
median_steps_final<-tapply(act_final$steps, days1, median)
median_steps_final #Medians for the days with missing values differed from 0 because the substitution was made with means
plot(median_steps_final)
```



## Are there differences in activity patterns between weekdays and weekends?

```{r, results="asis"}
days2<-as.POSIXlt(days1)
class(days2)
days3<-days2$wday
fdays<-factor(days3)
act_final$fdays<-fdays
colnames(act_final)
sub_wend<-subset(act_final, subset=(act_final$fdays%in%c(0,6)))
sub_wday<-subset(act_final, subset=(act_final$fdays%in%c(1:5)))
dim(sub_wend)
dim(sub_wday)
mean_sub_wday<-tapply(sub_wday$steps,sub_wday$interval,mean)
mean_sub_wend<-tapply(sub_wend$steps,sub_wend$interval,mean)
par(mfrow=c(2,1))
plot(mean_sub_wday~levels(factor(sub_wday$interval)), type="l", main="Activity in the weekdays", xlab="Time interval")
plot(mean_sub_wend~levels(factor(sub_wend$interval)), type="l", main="Activity in the weekends", xlab="Time interval")
```

That is the end.
